A boundary-integral framework to simulate \\ viscous erosion of a porous medium

### 2018-04-13 docs/methods/erosion_methods.tex
NM

\begin{abstract} 
We develop numerical methods to simulate the fluid-mechanical erosion of many bodies in a two-dimensional Stokes flow. The broad aim is to simulate the erosion of a porous medium (e.g.~ground water flow) with grain-scale resolution. Our fluid solver is based on a second-kind boundary integral formulation of the Stokes equations that is discretized with a spectrally-accurate Nystr\"om method and solved with fast-multipole-accelerated GMRES. To evolve the individual grains under shear-driven erosion, we regularize via curvature-penalization and solve the associated system using the {\thL} formulation. This framework allows numerically stable treatment of stiff terms, thus permitting large time steps. The overall accuracy of our method is spectral in space and second-order in time. The method is computationally efficient, with the fluid solver requiring $\mathcal{O}(N)$ operations per GMRES iteration, a mesh-independent number of GMRES iterations, and a one-time $\mathcal{O}(N^2)$ computation to compute the shear stress. We benchmark single-body results against analytical predictions. In our multi-body simulations, we find that erosion leads to spontaneous channelization between bodies in near contact. The channelization is associated with a dramatic reduction in the resistance of the porous medium, much more than would be expected from the reduction in grain size alone.
\end{abstract}

@@

Recent work in the high-Reynolds-number regime has highlighted the importance of a shape-flow feedback that occurs during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann, lopez2018cfd} and during related processes such as dissolution and melting~\cite{Huang2015, kondratiuk2015steadily, dallaston2015channelization, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}.  In these cases, the fluid alters the morphology of the immersed structures, which in turn modifies the surrounding flow field, and so on. Such examples of `sculpting with flow'~\cite{ristroph2018sculpting} have been examined in the context of a single body~\cite{ris-moo-chi-she-zha2012, moore2017riemann} and, more recently, multiple bodies whose shapes coevolve as mediated by the flow~\cite{hewett2017evolution}. From this viewpoint, the example of an eroding porous medium presents new opportunities and challenges. First, since the medium is composed of many individual constituents, multiple bodies must be considered. Second, the tight-packing present in typical porous-media applications implies that close-range interactions must be resolved. Finally, long-range effects inherent to Stokes flow couples the shape evolution of {\em all} bodies to a greater degree than any other flow regime.

@@

%Comparing the 50 and 6-body simualtions... Though the area fractions of the two simulations are nearly the same, the 50-body configuration provides about 5 times more resistance.

% Good word: Spontaneous channelization

>>>>

\begin{abstract} 
We develop numerical methods to simulate the fluid-mechanical erosion of many bodies in two-dimensional Stokes flow. The broad aim is to simulate the erosion of a porous medium (e.g.~ground water flow) with grain-scale resolution. Our fluid solver is based on a second-kind boundary integral formulation of the Stokes equations that is discretized with a spectrally-accurate Nystr\"om method and solved with fast-multipole-accelerated GMRES. To evolve the individual grains under shear-driven erosion, we regularize via curvature-penalization and solve the associated system using the {\thL} formulation. This framework allows numerically stable treatment of stiff terms, thus permitting large time steps. The overall accuracy of our method is spectral in space and second-order in time. The method is computationally efficient, with the fluid solver requiring $\mathcal{O}(N)$ operations per GMRES iteration, a mesh-independent number of GMRES iterations, and a one-time $\mathcal{O}(N^2)$ computation to compute the shear stress. We benchmark single-body results against analytical predictions. In our multi-body simulations, we find that erosion leads to spontaneous channelization between bodies in near contact. The channelization is associated with a dramatic reduction in the resistance of the porous medium, much more than would be expected from the reduction in grain size alone.
\end{abstract}

@@

Recent work in the high-Reynolds-number regime has highlighted the importance of a shape-flow feedback that occurs during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann, lopez2018cfd} and during related processes such as dissolution and melting~\cite{Huang2015, kondratiuk2015steadily, dallaston2015channelization, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}.  In these cases, the fluid alters the morphology of the immersed structures, which in turn modifies the surrounding flow field, and so on. Such examples of `sculpting with flow'~\cite{ristroph2018sculpting} have been examined in the context of a single body~\cite{ris-moo-chi-she-zha2012, moore2017riemann} and, more recently, multiple bodies whose shapes coevolve as mediated by the flow~\cite{hewett2017evolution}. From this viewpoint, the example of an eroding porous medium presents new opportunities and challenges. First, since the medium is composed of many individual constituents, multiple bodies must be considered. Second, the dense packing present in typical porous-media applications requires close-range interactions to be resolved. Finally, long-range effects inherent to Stokes flow couples the shape evolution of {\em all} bodies to a greater degree than any other flow regime.

@@
